%
%                              POK header
% 
% The following file is a part of the POK project. Any modification should
% be made according to the POK licence. You CANNOT use this file or a part
% of a file for your own project.
%
% For more information on the POK licence, please see our LICENCE FILE
%
% Please follow the coding guidelines described in doc/CODING_GUIDELINES
%
%                                      Copyright (c) 2007-2009 POK team 
%
% Created by julien on Thu Apr 30 14:07:20 2009 
%


\chapter{Installation}
   \label{chapter-installation}
   \section{Supported development platforms}
   \begin{itemize}
      \item[$\bullet$] \linux
      \item[$\bullet$] \macosx
      \item[$\bullet$] \windows
   \end{itemize}

   \section{Get more information}
   The following information are the standard procedures. It may be out of date
   or miss something. In that case, you will find updated information on the
   \pok website (\texttt{http://pok.gunnm.org}) and its wiki section.

   In addition, there are some tutorials and information about the installation
   of required tools.

   \section{Linux/MacOS}
      \subsection{Pre-requires}
      \begin{itemize}
         \item[$\bullet$] The GNU-C Compiler (aka GCC), version 3.x or 4.x
         \item[$\bullet$] GNU binutils
         \item[$\bullet$] GNU Zip (aka gzip)
         \item[$\bullet$] Mtools (MS-DOS disk utilities)
         \item[$\bullet$] AWK
         \item[$\bullet$] Perl (with \texttt{XML::XPath::XMLParser} and
            \texttt{XML::LibXML} modules)
         \item[$\bullet$] \qemu (for x86 emulation)
         \item[$\bullet$] \ocarina (for code generation only)
         \item[$\bullet$] TSIM (for Leon3 emulation)
      \end{itemize}

      \subsubsection{Note for MacOS users}
      \pok uses the ELF format to store partitions. Unfortunately, this binary
      format is not supported by \macosx tools. To use \pok, you must use a
      development toolchain that supports the ELF format.

      For that, you can easily build an ELF cross compiler using MacPorts. The
      name of the required packages are \texttt{i386-elf-gcc} and
      \texttt{i386-elf-binutils}.

      Moreover, \macosx does not provide necessary Perl modules but you can
      install them with MacPorts. The package names are \texttt{p5-xml-xpath} and
      \texttt{p5-xml-libxml}.


\subsection{Running POK}

Ocarina is needed by POK. A script is provided to automatically install
the latest build:

\begin{verbatim}
$ sh ./misc/get_ocarina.sh
\end{verbatim}

You can then try to build and run some of the POK examples located in the
`example' directory.

\begin{verbatim}
$ cd examples/partitions-threads
$ make
$ make -C generated-code run
\end{verbatim}

A whole chapter of this documentation is dedicated to those examples and their
purpose.


   \section{Windows}
      \subsection{Pre-requires}
      There is many pre-requires to run \pok on Windows. To make a better user
      experience, we provide cross-development tools to build and use POK.

      For cross-development tools can be retrieved from the website of the
      project. Then, unzip the tools in a directory called \texttt{crosstools},
      at the root directory of the project.

      Once you have this directory, run the file \texttt{configure.bat} located
      in this directory. Is everything is wrong, a warning will be displayed on
      the screen.

      For code generation, you can install \ocarina for Windows. All
      installation instructions are available on Ocarina website.

